---
title: "Experiment 2 - social bond - power simulation"
author: "Lucrezia Lonardo"
date: "14/05/2025"
output: 
  html_document:
    theme: united
    toc: yes
    toc_depth: 4
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
library(tidyverse)
library(cowplot)
library("gghalves")
library(ggthemes)
library(betareg)
library(lmtest)
source("../functions/beta_par_to_par_transf.r")

#load("social_bond_exp2_power_sim_N64.RData")

```

## Generate data

```{r echo=FALSE, include=FALSE}
set.seed(1)
n.subject <- 64# number subjects (nr. of children tested in Wolf & Tomasello, 2020)
n <- 128 #nrow(final dataset)=tot. nr. obs
n.per.subject <- 2 # observations per subject
n.per.condition <- 1 # observations per subject and condition
subj.id <- as.factor(paste("subj", str_pad(1:n.subject, 2, pad = "0"), sep = "."))

test.per <- c(0.19) # performance in test condition
control.per <- c(0.30, 0.37) # performance in control condition
```


```{r echo=FALSE, include=FALSE}
# LL: OLD CODE
# start.data <- data.frame(subj.id)
# xdata=data.frame(expand.grid(condition=c("control", "test"), sex=c("f", "m")))
# start.data<-cbind(start.data, xdata)

# Split subjects into two groups for counterbalancing
half <- length(subj.id) / 2
first_half <- subj.id[1:half]
second_half <- subj.id[(half + 1):length(subj.id)]
sex.per.subj <- rep(c("f", "m"), each = n.subject / 2)

# Create trial data for each subject
start.data <- data.frame()

for (i in seq_along(subj.id)) {
  s <- subj.id[i]
  
  # Half get control first, half get test first
  conditions <- if (i <= n.subject / 2) c("control", "test") else c("test", "control")
  
  subj_data <- data.frame(
    subj.id = s,
    trial = 1:2,
    condition = conditions
   # ,sex = sex.per.subj[s]
  )
  
  start.data <- rbind(start.data, subj_data)
}

start.data$sex=rep(c("f", "m"), each=2)
table(start.data$condition, start.data$sex)
table(start.data$subj.id, start.data$sex)

start.data$order<- rep(c("control_first", "test_first"), each=n.subject)

# z-transformation of trial nr.
start.data$z.trial <- as.vector(scale(as.numeric(start.data$trial)))
```


## Simulation

```{r eval=FALSE, include=FALSE}
n.simus <- 1000 
resid.sd <- 350#residual standard deviation
subj.sd <- 500#random effect of individual

icpt <- control.per#intercept
sex.effect <- 0
order.effect <- 0

# create object to store the simulation parameters and results:
all.res <- data.frame(expand.grid(
  n.per.subject = n.per.subject, 
  test.per = test.per,
  control.per = control.per,
  simu = 1:n.simus
))
all.res$icpt <- NA
all.res$conditiontest <- NA
#added by LL (needed?)
all.res$re.sd <- NA
all.res$lrt.p.con <- NA
all.res$lrt.p.con <- NA
all.res$lrt.p.sex <- NA
all.res$lrt.p.order <- NA

all.res$phi <- NA
all.res$warns.full <- NA
all.res$warns.null <- NA
all.res$lrt.p.con <- NA
all.res$full.null.p <- NA

all.ests <- matrix(NA, nrow = n.simus, ncol = 1)
colnames(all.ests) <- c("lrt.p.con")

# create data frame with design:
## done above

# load packages needed:
library(lme4)
# Loading required package: Matrix
library(kyotil) # we want to store info about convergence issues

# define control structure to make convergence more likely:
xdata <- start.data
str(xdata)
xdata$condition<-as.factor(xdata$condition)
xdata$order<-as.factor(xdata$order)
xdata$sex<-as.factor(xdata$sex)
```


```{r eval=FALSE, include=FALSE}
# run simulation
for (i in 1:nrow(all.res)) {
  set.seed(i) # allows to later replicate individual simulations

  m.mat <-
    model.matrix(object = ~ condition + sex + order, data = xdata) # create model martix
  
  coefs <- c(
    "(Intercept)" = qlogis(all.res[i, "control.per"]),
    #performance in control condition
    "conditiontest" = qlogis(all.res[i, "test.per"]) - qlogis(all.res[i, "control.per"]),
    #performance in test condition
    "sexm" = 0,
    "ordertest_first" = 0
  )
  
  LP <- m.mat[, names(coefs)] %*% coefs # LP wrt fixed effects
 
  #LL added 
  # add random intercept of subj
  library(MASS)  # for mvrnorm

# Number of subjects
n.subj <- length(unique(xdata$subj.id))

# Variance-covariance components
intercept.var <- 0.36       # SD = 0.6
slope.var <- 0.16           # SD = 0.4
cor.is <- 0.2               # correlation between intercept & slope
cov.is <- cor.is * sqrt(intercept.var) * sqrt(slope.var)

# Create covariance matrix
Sigma <- matrix(c(
  intercept.var, cov.is,
  cov.is,        slope.var
), ncol = 2)

# Simulate random effects
re.mat <- MASS::mvrnorm(n = n.subj, mu = c(0, 0), Sigma = Sigma)

# Extract subject-specific intercepts and slopes
subj.int <- re.mat[, 1]
subj.slope <- re.mat[, 2]

# Add to the linear predictor
cond <- model.matrix(~ condition, data = xdata)[, "conditiontest"]

LP <- LP + subj.int[as.numeric(xdata$subj.id)] +
          subj.slope[as.numeric(xdata$subj.id)] * cond

  
  phi = 10##precision parameter
  yvals = exp(LP) / (1 + exp(LP))#proportion
  #transform mean/var to shape1/shape2 parameters:
  beta.par = beta.par1topar2(mu = yvals, phi = phi)
  #simulate/generate response:
  prop.bc = rbeta(n = n,
                  shape1 = beta.par$shape1,
                  shape2 = beta.par$shape2)
 prop.bc.scaled <- (prop.bc*(length(prop.bc) - 1) + 0.5)/length(prop.bc)#transform DV to exclude 0 and 1s

  #LL:old code
  # # fit full model:
  # full <- keepWarnings(betareg(prop.bc ~ condition + sex + order,
  #                              data = xdata))
  # # fit null model:
  # null <- keepWarnings(betareg(prop.bc ~ sex + order,
  #                              data = xdata))
   ##LL added
  library(glmmTMB)
# fit full model:
full <- glmmTMB(
  prop.bc.scaled ~ condition + sex + order + (1 | subj.id),
  family = beta_family(),
  data = xdata
)
# fit null model:
null <- glmmTMB(
  prop.bc.scaled ~ sex + order + (1 | subj.id),
  family = beta_family(),
  data = xdata
)
  
coefs_full <- fixef(full)$cond
  
  # Store results
# Extract fixed effects
coefs_full <- fixef(full)$cond

# Store coefficients and dispersion (phi)
all.res[i, c("icpt", "conditiontest", "sexm", "ordertest_first")] <- coefs_full
all.res[i, "phi"] <- sigma(full)

# Likelihood ratio test for effect of condition
lrt.condition <- anova(full, null)
all.res[i, "lrt.p.con"] <- lrt.condition$"Pr(>Chisq)"[2]

# Track convergence
all.res[i, "converged"] <- ifelse(isTRUE(all.equal(full$sdr$pdHess, TRUE)), 1, 0)

}

save.image("social_bond_exp2_power_sim_N64.RData")
```

```{r}
  #plot to check:
  par(mar=c(3, 3, 1.2, 0.2), mgp=c(1.7, 0.3, 0), tcl=-0.15, las=1)
  boxplot(prop.bc~xdata$condition)
```

## Evaluation of results 

## Only models that converged are evaluated from here on:  

```{r include=FALSE}
# Filter for successfully converged models
all.res2 <- all.res %>% filter(converged == 1) #all models converged
```


### How many models converged, have a significant full-null model comparison, and a significant LRT of condition?  
```{r echo=FALSE}
# Total number of converged models
n.converged <- nrow(all.res)


lrt.data2 <- all.res2 %>%
  filter(full.null.p<0.05)%>%
  group_by(test.per,control.per) %>%
  summarise(lrt.p.con.median = median(lrt.p.con),
            lrt.p.fullnull.median = median(full.null.p),
            n.sign.lrt.con = length(lrt.p.con[lrt.p.con < 0.05]),
            n.sign.lrt.fullnull = length(full.null.p[full.null.p < 0.05]),
            n.lrt = n.simus,
            proportion.sign.lrt.con = length(lrt.p.con[lrt.p.con < 0.05]) / n.simus,
            proportion.sign.lrt.fullnull = length(full.null.p[full.null.p < 0.05]) / n.simus)
lrt.data2
```
#### Plotting the proportion of significant LRTs for the predictor variable condition ONLY based on models that converged and with a significant full-null model comparison

```{r echo=FALSE}
p.con.power <- ggplot(data = lrt.data2, aes(x= as.factor(test.per),y = proportion.sign.lrt.con, fill=as.factor(control.per))) +
  geom_bar(stat="identity", color="black", position=position_dodge())+
  scale_y_continuous(breaks=seq(0,1,0.2), limits=c(0, 1))+
  geom_hline(yintercept = 0.8, colour = "black", lwd = 1.1, lty = 2) +
    geom_hline(yintercept = 0.05, colour = "darkgrey", lwd = 1.1, lty = 4) +
  scale_fill_manual(values=c("dodgerblue", "darkorange"))+
  labs(fill = "control condition", y="Power", x= "test condition") +
  theme_few()#+
  #theme(legend.position="none")
p.con.power

ggsave(p.con.power, filename = "./exp2_beta_glm.png", scale = 0.8, height = 5, width = 7)
```