---
title: "Study 1 - dog-human social bond - power simulation"
author: "Lucrezia Lonardo"
date: "14/05/2025"
output: 
  html_document:
    theme: united
    toc: yes
    toc_depth: 4
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
library("gghalves")

#load("dog_human_social_bond_exp1_power_sim_N36.RData")
```

## Generate data

```{r echo=FALSE, include=FALSE}
n.subject <- 36 # number subjects (currently ET-trained dogs)
n.per.subject <- 8 # observations per subject
n.per.condition <- 4 # observations per subject and condition
subj.id <- as.factor(paste("subj", str_pad(1:n.subject, 2, pad = "0"), sep = "."))
constricting.per <- c(-250) # performance in constricting condition
dilating.per <- c(250) # performance in dilating condition
```


```{r echo=FALSE, include=FALSE}
# Assign counterbalancing order
# Even subjects get "constricting" first, odd get "dilating" first 
cb_order <- ifelse(as.numeric(subj.id) %% 2 == 0, "constricting", "dilating")

# Initialize empty data frame
start.data <- data.frame()

# Loop through each subject and create their trials
for (i in seq_along(subj.id)) {
  subj <- subj.id[i]
  first_cond <- cb_order[i]
  second_cond <- ifelse(first_cond == "constricting", "dilating", "constricting")
  
  conditions <- c(rep(first_cond, n.per.condition), rep(second_cond, n.per.condition))
  
  subj_data <- data.frame(
    subj.id = subj,
    condition = conditions,
    trial = 1:(2 * n.per.condition)
  )
  
  start.data <- rbind(start.data, subj_data)
}

# add order
start.data$order <- as.factor(rep(x = c("dilating_first", "constricting_first"), times = n.subject / 2))[as.numeric(start.data$subj.id)]

start.data$order2 <- ifelse((start.data$order == "dilating_first" & start.data$condition == "dilating") |
  (start.data$order == "constricting_first" & start.data$condition == "constricting"), 1, 2)


# add demographics 
start.data$sex <- as.factor(rep(x = c("f", "m", "m", "f"), times = n.subject/4))[as.numeric(start.data$subj.id)]


# z-transformation of covariates
start.data$z.order <- as.vector(scale(as.numeric(start.data$order2)))
```

# Data checks:

```{r echo=FALSE, include=FALSE}
# does each subject have only one sex and age?
xx <- table(start.data$subj.id, start.data$sex)
range(apply(X = xx > 0, MARGIN = 1, sum)) # should be 1 and 1

xx <- table(start.data$order, start.data$sex)
range(apply(X = xx > 0, MARGIN = 1, sum))

xx <- table(start.data$subj.id, start.data$condition)
range(apply(X = xx > 0, MARGIN = 1, sum))

xx <- table(start.data$condition, start.data$order)
range(apply(X = xx > 0, MARGIN = 1, sum))

xx <- table(start.data$subj.id, start.data$order)
range(apply(X = xx > 0, MARGIN = 1, sum))
```


## Simulation

Test
```{r eval=FALSE, include=FALSE}
n.simus <- 1000 # small number for testing
resid.sd <- 350#residual standard deviation
subj.sd <- 500#random effect of individual

icpt <- constricting.per#intercept
sex.effect <- 0
order.effect <- 0

xdata<-start.data
tot.n <- nrow(xdata)#total n
m.mat <- model.matrix(object=~condition+sex+order, data=xdata)
coefs <- c("(Intercept)"=icpt, "conditiondilating"=dilating.per,  "sexm"=sex.effect, "orderdilating_first"=order.effect)

rv <- m.mat[, names(coefs)]%*%coefs+#fixed effects
rnorm(n=n.subject, sd=subj.sd)[as.numeric(subj.id)]+#random effect of subj.
rnorm(n=tot.n, sd=resid.sd)#residual variation

xdata$condition <- factor(xdata$condition)

#plotting
par(mar=c(2.7, 2.7, 0.2, 0.2), mgp=c(1.5, 0.3, 0),
tcl=-0.15, las=1, cex.lab=0.7, cex.axis=0.5)
plot(xdata$condition, rv)

library(lme4)
full=lmer(rv~condition+sex+order+(1|subj.id), data=xdata, REML=F)
summary(full)
drop1(full, test="Chisq")
```

Actual simulation
```{r eval=FALSE, include=FALSE}
n.simus <- 1000 
tot.n <- nrow(xdata)#total n
resid.sd <- 350#residual standard deviation
subj.sd <- 500#random effect of individual

icpt <- constricting.per#intercept
sex.effect <- 0
order.effect <- 0

# create object to store the simulation parameters and results:
all.res <- data.frame(expand.grid(
  n.subject = n.subject, r.effect = subj.sd, resid.sd=resid.sd, 
  dilating.per = dilating.per, constricting.per = constricting.per,
  simu = 1:n.simus
))
all.res$icpt <- NA
all.res$conditiondilating <- NA
all.res$re.sd <- NA
all.res$warns.full <- NA
all.res$lrt.p.con <- NA
all.res$lrt.p.con <- NA
all.res$lrt.p.sex <- NA
all.res$lrt.p.order <- NA

all.ests <- matrix(NA, nrow=n.simus, ncol=1)
colnames(all.ests)=c("lrt.p.con")

# create data frame with design:
## done above

# load packages needed:
library(lme4)
# Loading required package: Matrix
library(kyotil) # we want to store info about convergence issues

# define control structure to make convergence more likely:
contr <- lmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 10000))

xdata <- start.data
```


```{r eval=FALSE, include=FALSE}
# run simulation
for (i in 1:nrow(all.res)) {
  set.seed(i) # allows to later replicate individual simulations
  
  m.mat <-
    model.matrix(object = ~ condition + sex + order, data = xdata) # create model martix
  
  coefs <- c(
    "(Intercept)" = constricting.per,
    "conditiondilating" = dilating.per,
    "sexm" = 0,
    "orderdilating_first" = 0
  )
  # generate response:
  xdata$rv <- m.mat[, names(coefs)] %*% coefs + #fixed effects
    rnorm(n = n.subject, sd = subj.sd)[as.numeric(subj.id)] + #random effect of subj.
    rnorm(n = tot.n, sd = resid.sd)#residual variation
  
  # fit full model:
  full <-
    keepWarnings(lmer(
      rv ~ condition + sex +  order + (1 | subj.id),
      data = xdata,
      REML = FALSE,
      control = contr
    ))

  # store results:
  all.res[i, c("(Intercept)",
               "conditiondilating",
               "sexm",
               "orderdilating_first")] <- fixef(full$value)
  all.res[i, "re.sd"] <-
    as.data.frame(summary(full$value)$varcor)[1, "sdcor"]
  all.res[i, "warns.full"] <-
    nchar(paste(full$warnings, collapse = ""))
  all.res[i, "lrt.p.con"] <-
    as.data.frame(drop1(full$value, test = "Chisq"))["condition", "Pr(Chi)"]
   all.res[i, "lrt.p.sex"] <-
    as.data.frame(drop1(full$value, test = "Chisq"))["sex", "Pr(Chi)"]
    all.res[i, "lrt.p.order"] <-
    as.data.frame(drop1(full$value, test = "Chisq"))["order", "Pr(Chi)"]
}

save.image("dog_human_social_bond_exp1_power_sim_N36.RData")
```

## Evaluation of results 

* number of warning per combinations of random effects (out of 1000 models per cell)  
Full model:  
```{r echo=FALSE}
#full model
tapply(X=all.res[, "warns.full"]>0, INDEX=all.res[, c("constricting.per", "dilating.per")],
FUN=sum)
#warning codes: 
#Warning: convergence code 3 from bobyqa: bobyqa -- a trust region step failed to reduce qWarning: convergence code 3 from bobyqa: bobyqa -- a trust region step failed to reduce qWarning: convergence code 3 from bobyqa: bobyqa -- a trust region step failed to reduce qboundary (singular) fit: see help('isSingular')
```

* plotting the estimates (all models)

```{r echo=FALSE}
par(mar=c(3, 3, 0.2, 0.2), mgp=c(1.7, 0.3, 0), tcl=-0.15, las=1)
plot(
  x = as.numeric(as.factor(rep(
    x = c("(Intercept)", "conditiondilating", "re.sd"),
    each = nrow(all.res)
  ))),
  y = unlist(all.res[, c("(Intercept)", "conditiondilating", "re.sd")]),
  pch = 19, col = grey(level = 0.2, alpha = 0.2),
  xaxt = "n", xlim = c(0.5, 3.5), ylab = "estimate", xlab = ""
)
mtext(text = c("(Intercept)", "conditiondilating", "re.sd"), side = 1, at = 1:3, line = 0.2)
```

## Only models that converged are evaluated from here on:  

```{r include=FALSE}
all.res2=subset(all.res, warns.full==0)
```


### How many models converged, have a significant full-null model comparison, and a significant LRT of condition?  
```{r echo=FALSE}

lrt.data2 <- all.res2 %>%
  group_by(dilating.per, constricting.per) %>%
  summarise(lrt.p.con.mean2 = mean(lrt.p.con), 
            n.sign.lrt2 = length(lrt.p.con[lrt.p.con < 0.05]), 
            n.lrt = n.simus,#length(lrt.p.con), 
            proportion.sign.lrt2.condition = length(lrt.p.con[lrt.p.con < 0.05]) / n.simus,
            proportion.sign.lrt2.sex = length(lrt.p.con[lrt.p.sex < 0.05]) / n.simus)

lrt.data2
```


